\vspace{-1em}Guidance, please! Towards a framework for RDF-based constraint languages.

### 2015-07-07 paper/paper 2/dc-2015.tex
2.3

The implementation model of the \emph{RDF-CV} (Figure \ref{fig:RDF-CV-implementation-model}) shows that
both \emph{property constraints} and \emph{class constraints} are \emph{simple constraints}. 
%For both \emph{property} and \emph{class constraints} a \emph{context class}, a list of \emph{classes}, the \emph{constraining element}, and the \emph{constraining value} can be stated. 
%Lists of \emph{left} and \emph{right properties} can only be specified for \emph{property constraints}.
\begin{figure}[H]
	\centering
		\includegraphics[width=0.80\textwidth]{images/RDF-CV-implementation-model.png}
	\caption{\emph{RDF Constraints Vocabulary (RDF-CV)} Implementation Model}
	\label{fig:RDF-CV-implementation-model}
\end{figure}
A \emph{simple constraint} holds for all individuals of their associated \emph{context class}.
The \emph{minimum qualified cardinality restriction} {\small\ms{Publication $\sqsubseteq$ $\geq$1 author.Person,}}
which restricts \emph{publications} to have at least one \emph{author} which must be a \emph{person},
is an example of a \emph{property constraint} on \emph{author}
which holds for all individuals of the class \emph{Publication}.
Table \ref{tab:property-constraint-cardinality-restriction} displays how the \emph{property constraint} is generically represented using the \emph{RDF-CV}.
\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Minimum Qualified Cardinality Restriction as Property Constraint}
	\label{tab:property-constraint-cardinality-restriction}
	\centering
		\begin{tabular}{c|c|c|c|c|c|c}
      \textbf{constraint set} & \textbf{context class} & \textbf{left property list} & \textbf{right p. list} & \textbf{classes} & \textbf{constraining element} & \textbf{c. value} \\
      \hline
      property & Publication & author & - & Person & $\geq$ & 1 \\
		\end{tabular}
\end{table}

@@

In some cases, a constraint is only complete when a \emph{constraining value} is stated in addition to the \emph{constraining element}
and \emph{simple constraints} may refer to a list of \emph{classes}.
The \emph{constraining element} of the \emph{property constraint} {\small\ms{Publication $\sqsubseteq$ $\geq$1 author.Person},} e.g.,  is the \emph{DL at-least restriction} \emph{$\geq$}, the \emph{constraining value} is \emph{1}, and
the list of \emph{classes} includes the class \emph{Person} which restricts the objects of the property \emph{author} to be persons.

For \emph{property constraints}, \emph{left} and \emph{right property lists} are specified.
The assignment of properties to these lists happens relative to the \emph{constraining element} 
which may be a DL operator like $\sqsubseteq$.
\emph{Object Property Paths} (\emph{R-55})
ensure that if an individual \emph{x} is connected by a sequence of object properties with an individual \emph{y}, 
then \emph{x} is also related to \emph{y} by a particular object property. 
As \emph{Stephen-Hawking} is the author of the book \emph{A-Brief-History-Of-Time} 
%{\small\ms{(authorOf(Stephen-Hawking, A-Brief-History-Of-Time))}} 
whose genre is \emph{Popular-Science}, 
%{\small\ms{(genre(A-Brief-History-Of-Time, Popular-Science)),}}
the object property path {\small\ms{authorOf $\circ$ genre $\sqsubseteq$ authorOfGenre}} infers that \emph{Stephen-Hawking} is an author of the genre \emph{Popular-Science}. 
%{\small\ms{(authorOfGenre(Stephen-Hawking, Popular-Science))}.}
Thus, when representing the \emph{property constraint} using the \emph{RDF-CV} (see Table \ref{tab:property-constraint-object-property-paths}), the properties \emph{authorOf} and \emph{genre} are placed on the left side of the \emph{constraining element} \emph{$\sqsubseteq$}
and the property \emph{authorOfGenre} on its right side.
As this \emph{property constraint} holds for all individuals within the data, the \emph{context class} is set to the \emph{DL top concept} $\top$ which stands for the super-class of all classes.
\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Object Property Paths as Property Constraint}
	\label{tab:property-constraint-object-property-paths}
	\centering
		\begin{tabular}{l|l|l|l|l|l|l}
      \textbf{c. set} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
      property & $\top$ & authorOf, genre & authorOfGenre & $\top$ & $\sqsubseteq$ & - \\
		\end{tabular}
\end{table}

There are \emph{simple constraints} which are not expressible in DL but can still be described using the \emph{RDF-CV} 
such as constraints of the type \emph{literal pattern matching} (\emph{R-44}) which restrict literals to match given patterns.
The \emph{universal restriction} {\small\ms{Book $\sqsubseteq$ $\forall$ identifier.ISBN}} ensures that \emph{books} can only have valid \emph{ISBN} identifiers, 
i.e., strings that match a given regular expression.

@@

Table \ref{tab:simple-constraint-not-expressible-in-dl)} presents 
(1) how the \emph{literal pattern matching} \emph{simple constraint} which is not expressible in DL and 
(2) how the \emph{universal restriction} \emph{complex constraint} which is expressible in DL are represented using the \emph{RDF-CV}.
Thereby, the \emph{context class} \emph{ISBN}, whose instances must satisfy the \emph{simple constraint}, 
is reused within the list of \emph{classes} the \emph{complex constraint} refers to.
The \emph{literal pattern matching} constraint type introduces the \emph{constraining element} \emph{regex}
whose validation has to be implemented once like for any other \emph{constraining element}.

@@

      property & Book & identifier & - & ISBN & $\forall$ & - \\

@@

\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Simplified Complex Constraints}
	\label{tab:simplified-complex-constraints}
	\centering
		\begin{tabular}{l|l|l|l|l|l|l}
      \textbf{c. set} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
class & Publication & - & - & E, F & XOR \\
class & E & - & - & A, B & $\sqcap$ \\
class & F & - & - & C, D & $\sqcap$ \\
property & A & isbn & - & string & = & 1 \\
property & B & title & - & string & = & 1 \\
property & C & issn & - & string & = & 1 \\
property & D & title & - & string & = & 1 \\
		\end{tabular}
\end{table}

@@

\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Primary Key Properties as Complex Constraints}
	\label{tab:primary-key-properties-as-complex-constraints}
	\centering
		\begin{tabular}{l|l|l|l|l|l|l}
      \textbf{c. set} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
property & $\top$ & isbn$^{-}$ & isbn$^{-}$ & - & inverse & - \\
property & Book & isbn$^{-}$ & - & - & functional & - \\
		\end{tabular}
\end{table}

>>>>

The implementation model of the \emph{RDF-CV} (Figure \ref{fig:RDF-CV-implementation-model}) shows that
both \emph{property constraints} and \emph{class constraints} are \emph{simple constraints}. 
%For both \emph{property} and \emph{class constraints} a \emph{context class}, a list of \emph{classes}, the \emph{constraining element}, and the \emph{constraining value} can be stated. 
%Lists of \emph{left} and \emph{right properties} can only be specified for \emph{property constraints}.
\begin{figure}[H]
	\centering
		\includegraphics[width=0.80\textwidth]{images/RDF-CV-implementation-model.png}
	\caption{\emph{RDF Constraints Vocabulary (RDF-CV)} Implementation Model}
	\label{fig:RDF-CV-implementation-model}
\end{figure}
A simple constraint holds for all individuals of their associated \emph{context class}.
The \emph{minimum qualified cardinality restriction (R-75)} {\small\ms{Publication $\sqsubseteq$ $\geq$1 author.Person,}}
which restricts publications to have at least one author which must be a person,
is an example of a property constraint on \emph{author}
which holds for all individuals of the class \emph{Publication}.
Table \ref{tab:property-constraint-cardinality-restriction} displays how the property constraint is generically represented using the RDF-CV.
\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Minimum Qualified Cardinality Restriction as Property Constraint}
	\label{tab:property-constraint-cardinality-restriction}
	\centering
		\begin{tabular}{c|c|c|c|c|c|c}
      \textbf{constraint set} & \textbf{context class} & \textbf{left property list} & \textbf{right p. list} & \textbf{classes} & \textbf{constraining element} & \textbf{c. value} \\
      \hline
      property & Publication & author & - & Person & $\geq$ & 1 \\
		\end{tabular}
\end{table}

@@

In some cases, a constraint is only complete when a \emph{constraining value} is stated in addition to the constraining element
and simple constraints may refer to a list of \emph{classes}. The constraining element of the property constraint {\small\ms{Publication $\sqsubseteq$ $\geq$1 author.Person},} e.g., is \emph{$\geq$}, the constraining value is \emph{1}, and the list of classes includes the class \emph{Person} which restricts the objects of the property \emph{author} to be persons.

For property constraints, \emph{left} and \emph{right property lists} are specified.
The assignment of properties to these lists happens relative to the constraining element.
\emph{Object Property Paths} (\emph{R-55})
ensure that if an individual \emph{x} is connected by a sequence of object properties with an individual \emph{y}, 
then \emph{x} is also related to \emph{y} by a particular object property. 
As \emph{Stephen-Hawking} is the author of the book \emph{A-Brief-History-Of-Time} 
%{\small\ms{(authorOf(Stephen-Hawking, A-Brief-History-Of-Time))}} 
whose genre is \emph{Popular-Science}, 
%{\small\ms{(genre(A-Brief-History-Of-Time, Popular-Science)),}}
the object property path {\small\ms{authorOf $\circ$ genre $\sqsubseteq$ authorOfGenre}} infers that \emph{Stephen-Hawking} is an author of the genre \emph{Popular-Science}. 
%{\small\ms{(authorOfGenre(Stephen-Hawking, Popular-Science))}.}
Thus, when representing the property constraint using the RDF-CV (see Table \ref{tab:property-constraint-object-property-paths}), the properties \emph{authorOf} and \emph{genre} are placed on the left side of the constraining element \emph{property paths}
and the property \emph{authorOfGenre} on its right side. As this property constraint holds for all individuals within the data, the context class is set to the \emph{DL top concept} $\top$ which stands for the super-class of all possible classes.
\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Object Property Paths as Property Constraint}
	\label{tab:property-constraint-object-property-paths}
	\centering
		\begin{tabular}{l|l|l|l|l|l|l}
      \textbf{c. set} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
      property & $\top$ & authorOf, genre & authorOfGenre & $\top$ & property paths & - \\
		\end{tabular}
\end{table}

There are simple constraints which are not expressible in DL but can still be described using the RDF-CV such as constraints of the type \emph{literal pattern matching} (\emph{R-44}) which restrict literals to match given patterns. The \emph{universal quantification (R-91)} {\small\ms{Book $\sqsubseteq$ $\forall$ identifier.ISBN}} ensures that \emph{books} can only have valid \emph{ISBN} identifiers, i.e., strings that match a given regular expression.

@@

Table \ref{tab:simple-constraint-not-expressible-in-dl)} presents (1) in the first line how the literal pattern matching simple constraint which is not expressible in DL and (2) in the second line how the universal quantification complex constraint which is expressible in DL are represented using the RDF-CV. Thereby, the context class \emph{ISBN}, whose instances must satisfy the simple constraint, is reused within the list of classes the complex constraint refers to. The literal pattern matching constraint type introduces the constraining element \emph{regex} whose validation has to be implemented once like for any other constraining element.

@@

      property & Book & identifier & - & ISBN & universal quantification & - \\

@@

\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Simplified Complex Constraints}
	\label{tab:simplified-complex-constraints}
	\centering
		\begin{tabular}{l|l|l|l|l|l|l}
      \textbf{c. set} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
class & Publication & - & - & E, F & XOR \\
class & E & - & - & A, B & intersection \\
class & F & - & - & C, D & intersection \\
property & A & isbn & - & string & = & 1 \\
property & B & title & - & string & = & 1 \\
property & C & issn & - & string & = & 1 \\
property & D & title & - & string & = & 1 \\
		\end{tabular}
\end{table}

@@

\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Primary Key Properties as Complex Constraints}
	\label{tab:primary-key-properties-as-complex-constraints}
	\centering
		\begin{tabular}{l|l|l|l|l|l|l}
      \textbf{c. set} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
property & $\top$ & isbn$^{-}$ & isbn$^{-}$ & - & inverse property & - \\
property & Book & isbn$^{-}$ & - & - & functional property & - \\
		\end{tabular}
\end{table}
